A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-10-06 inst/paper.Rnw
intro proof-read

\hh{todo list:
\begin{enumerate}
\item change format to comply with journal
\item include installation instruction for package ggnet
\item the biggest change to the previous implementation is that there is no longer a vertices data set. Instead, all of the vertex information has to be included in the edge dataset. The assumption is that any vertex information included relates to the `from\_id' column. If the user does not include ALL vertex information in the dataset (i.e. if some vertices only show up under the `to\_id' column), there will be segment ends without a dot in the graphic. As a measure against that, we suggest to merge edges and vertex information along the `from\_id' column using the parameter setting `all = TRUE'. This ensures all vertices to be included; if a vertex is a sink, the corresponding `to\_id' value is set to NA. This is accommodated for both in the layout and the drawing of the graph.
\item the parameters have changed names: all of the vertex parameters have the regular names, edge parameters start with `e'. `esize' is re-named to `linewidth'.
Some parameters might not be implemented yet, we need to get a good overview of what is missing and what works.
\item we could draw selfies, but at the moment we don't. We need to discuss this.
\end{enumerate}
}
\tableofcontents
\newpage
\section{Introduction}
At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.  Here, we refer to the lines as edges and the points as vertices, although these are also called nodes.  These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study. For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.  Additionally, \citet{networkfailures} examine node failure in interdependent networks like power grids.  Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C. elegans} worm are also etensively studied \citep{smallworld}.  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used and most complex networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.  The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network. The world wide web is an example of a directed network because one webpage may link to another, and not necessarily the other way around. Undirected edges, however, are simply connections between vertices.  Coauthorship networks that encode information about academic publications are examples of undirected networks because if two people author a paper together, that creates a connection between them that is bidirectional.

>>>>

\hh{todo list:
\begin{enumerate}
\item change format to comply with journal
\item include installation instruction for package ggnet
\item the biggest change to the previous implementation is that there is no longer a vertices data set. Instead, all of the vertex information has to be included in the edge dataset. The assumption is that any vertex information included relates to the `from\_id' column. If the user does not include ALL vertex information in the dataset (i.e. if some vertices only show up under the `to\_id' column), there will be segment ends without a dot in the graphic. As a measure against that, we suggest to merge edges and vertex information along the `from\_id' column using the parameter setting `all = TRUE'. This ensures all vertices to be included; if a vertex is a sink, the corresponding `to\_id' value is set to NA. This is accommodated for both in the layout and the drawing of the graph.
\item the parameters have changed names: all of the vertex parameters have the regular names, edge parameters start with `e'. `esize' is re-named to `linewidth'.
Some parameters might not be implemented yet, we need to get a good overview of what is missing and what works.
\item we could draw selfies, but at the moment we don't. We need to discuss this.
\end{enumerate}
}
\tableofcontents
\newpage
\section{Introduction}
At its core, a network is simply a set of points connected in pairs by a set of lines \citep{newman}.  Here, we refer to the lines as edges and the points as vertices, although these are also called nodes.  These two seemingly simple sets of graphical objects, points and segments, are used to encode a huge variety and quantity of information across many fields of study. For instance, networks of scientific collaboration, a food web of marine animals, and American college football games are all covered in a paper on community detection in networks by \citet{football}.  \citet{networkfailures} examine node failure in interdependent networks like power grids.  Social networks, such as links between actors found on \url{www.imdb.com}, and neural networks, like the completely mapped neural network of the \textit{C. elegans} worm are also etensively studied \citep{smallworld}.  Networks vary widely in scope and complexity: the smallest network is simply an edge between two vertices, while one of the most commonly used and most complex networks, the world wide web, has billions of vertices (webpages) and billions of edges (hyperlinks) connecting them.  The edges in a network can  be directed or undirected: directed edges represent information travelling from one vertex to another, and switching the direction would change the structure of the network. The world wide web is an example of a directed network because one webpage may link to another, but not necessarily the other way around. Undirected edges are simply connections between vertices. In co-authorship networks nodes are authors connected by an edge, if they author an academic publication together. Co-authorship networks are examples of undirected networks because if two people author a paper together, it creates a connection between them that is bidirectional.
